Cover Times of Random Walks

### 2019-05-07 talk.tex
minor

@@

		\node (m) at (-1.2,1.75) {
			\begin{tikzpicture}[scale = 3]
			    \tikzstyle{edge}=[draw,line width = .5pt,-,black!100]
			    \tikzstyle{current}=[draw,circle,fill=black,scale=.5]
			    \tikzstyle{visited}=[draw,circle,fill=white,scale=.5]
			    \draw[edge] (0,0) -- (0,1);
			    \draw[edge] (0,0) -- (1,0);
			    \draw[edge] (0,1) -- (1,0);
			    \draw[edge] (0,1) -- (1,1);
			    \draw[edge] (1,1) -- (1,0);
			    \node (l) at (.5, 1.1) {$t_{cov}$};
			    \node (bl) at (0,0) {};
			    \node[current] (tl) at (0,1) {};
			    \node (br) at (1,0) {};
			    \node (tr) at (1,1) {};;
			\end{tikzpicture}
		};
		\pause

@@

\begin{frame}{A Small Graph}
\small
\begin{align}
\E[t_{cov}] &= 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{b} 
\only<2->{= \frac{43}{6}}\nonumber \\
\textrm{a} &= 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{c} \nonumber \\
\textrm{b} &= 1 + \frac{1}{2}\textrm{b} + \frac{1}{2}\textrm{e} \nonumber \\
\textrm{c} &= 1 + 1\textrm{e} \nonumber \\
\textrm{d} &= 1 + \frac{1}{3}\textrm{a} + \frac{1}{3}\textrm{e} + \frac{1}{3}\textrm{f} \nonumber \\
\textrm{e} &= 1 + \frac{1}{3}\textrm{c} + \frac{1}{3}\textrm{e} + \frac{1}{3}\dot 0\nonumber \\
\textrm{f} &= 1 + \frac{1}{2}\textrm{g} + \frac{1}{2}\dot 0 \nonumber \\
\textrm{g} &= 1 + \frac{2}{3}\textrm{f} + \frac{1}{3}\dot 0\nonumber
\end{align}
\end{frame}

@@

\begin{frame}{Matthews Method Proof}
\only<1-3>{Let $\sigma$ be a uniform permutation of unvisited vertices from (worst-case) vertex $n$.}

\only<2-3>{Let $T_k$ be the first time vertices 
$\sigma_1, \sigma_2, \ldots, \sigma_k$ have all been visited.}
\only<3>{
\begin{align}
\E[t_{cov}] &= \E_n[{T_{n-1}}]  \nonumber \\
&= \E_n[T_1 + (T_2 - T_1) + \cdots + (T_{n-1} - T_{n-2})] . \nonumber
\end{align}
}

\only<4-> {Let $L(k)$ be the final vertex to be visited among 
$\sigma_1, \sigma_2, \ldots, \sigma_k$ before the first time they are
all visited.}
\only<5-> {
\begin{align}
\E_n[T_k - T_{k-1}] &=
\sum_{i=1}^k \E_n[T_k - T_{k-1} | L(k) = \sigma_i]
\pr(L_k = \sigma_i). \nonumber
\only<6->{\\&= 0 +
\E_n[T_k - T_{k-1} | L(k) = \sigma_k]
\pr(L_k = \sigma_k). \nonumber 
}
\only<7->{\\&\leq t_{hit} \frac{1}{k} \nonumber }
\end{align}
}
\end{frame}

@@

\begin{frame}{Poisson Distribution}
Let the number of unvisited vertices be $r = n \log n + nx$ (trust me).
\begin{align}
N_n &= ne^{-r/n} = ne^{-\log n - x} = n n^{-1} e^{-x} \rightarrow  e^{-x} \nonumber
\end{align}
\begin{theorem}
If $ne^{-r/n} \rightarrow \lambda \in [0, \infty)$, then the number of unvisited
vertices approaches a Poisson distribution with mean $\lambda$.
\end{theorem}
\end{frame}

\begin{frame}{Poisson Distribution}
Uppertail of a Poisson with mean $\lambda$
\begin{align}
\pr(N_n > 0) &\rightarrow 1 - e^\lambda \nonumber
\end{align}
\pause
The first time $N_n=0$ is $t_{cov}$
\begin{align}
\pr(t_{cov} \leq r) &\rightarrow e^{-e^{-x}} \nonumber \\
\pr(t_{cov} - n \log n\leq nx) &\rightarrow e^{-e^{-x}} \nonumber
\end{align}
\end{frame}

\begin{frame}{Birthday Problem}

@@

>>>>

\begin{frame}{Overview}
\tableofcontents
\end{frame}

@@

		\node (m) at (-1.2,1.75) {
			\begin{tikzpicture}[scale = 3]
			    \tikzstyle{edge}=[draw,line width = .5pt,-,black!100]
			    \tikzstyle{current}=[draw,circle,fill=black,scale=.5]
			    \tikzstyle{visited}=[draw,circle,fill=white,scale=.5]
			    \draw[edge] (0,0) -- (0,1);
			    \draw[edge] (0,0) -- (1,0);
			    \draw[edge] (0,1) -- (1,0);
			    \draw[edge] (0,1) -- (1,1);
			    \draw[edge] (1,1) -- (1,0);
%			    \node (l) at (.5, 1.1) {$t_{cov}$};
			    \node (bl) at (0,0) {};
			    \node[current] (tl) at (0,1) {};
			    \node (br) at (1,0) {};
			    \node (tr) at (1,1) {};;
			\end{tikzpicture}
		};
		\pause

@@

\begin{frame}{A Small Graph}
\small
\begin{align}
\E[t_{cov}] &= 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{b} 
\only<2->{= \frac{43}{6}}\nonumber \\
\textrm{a} &= 1 + \frac{1}{3}\textrm{a} + \frac{2}{3}\textrm{c} \nonumber \\
\textrm{b} &= 1 + \frac{1}{2}\textrm{b} + \frac{1}{2}\textrm{e} \nonumber \\
\textrm{c} &= 1 + 1\textrm{e} \nonumber \\
\textrm{d} &= 1 + \frac{1}{3}\textrm{a} + \frac{1}{3}\textrm{e} + \frac{1}{3}\textrm{f} \nonumber \\
\textrm{e} &= 1 + \frac{1}{3}\textrm{c} + \frac{1}{3}\textrm{e} + \frac{1}{3} 0\nonumber \\
\textrm{f} &= 1 + \frac{1}{2}\textrm{g} + \frac{1}{2} 0 \nonumber \\
\textrm{g} &= 1 + \frac{2}{3}\textrm{f} + \frac{1}{3} 0\nonumber
\end{align}
\end{frame}

@@

\begin{frame}{Matthews Method Proof}
Let $\sigma$ be a uniform permutation of unvisited vertices from (worst-case) vertex $n$.
\pause

Let $T_k$ be the first time vertices 
$\sigma_1, \sigma_2, \ldots, \sigma_k$ have all been visited.
\pause
\begin{align}
\E[t_{cov}] &\leq \E_n[{T_{n-1}}]  \nonumber \\
&= \E_n[T_1 + (T_2 - T_1) + \cdots + (T_{n-1} - T_{n-2})] . \nonumber
\end{align}

\end{frame}
\begin{frame}{Matthews Method Proof}

Let $L(k)$ be the final vertex to be visited among 
$\sigma_1, \sigma_2, \ldots, \sigma_k$ before the first time they are
all visited.
\pause
\begin{align}
\E_n[T_k - T_{k-1}] &=
\sum_{i=1}^k \E_n[T_k - T_{k-1} | L(k) = \sigma_i]
\pr(L_k = \sigma_i). \nonumber
\only<3->{\\&= 0 +
\E_n[T_k - T_{k-1} | L(k) = \sigma_k]
\pr(L_k = \sigma_k). \nonumber 
}
\only<4->{\\&\leq t_{hit} \frac{1}{k} \nonumber }
\end{align}
\end{frame}

@@

\begin{frame}{Poisson Distribution}
\begin{align}
r &= n \log n + nx \nonumber \\
N_n &= ne^{-r/n} = ne^{-\log n - x} = n n^{-1} e^{-x} \rightarrow  e^{-x} \nonumber
\end{align}
\begin{theorem}
If $ne^{-r/n} \rightarrow \lambda \in [0, \infty)$, then the number of unvisited
vertices approaches a Poisson distribution with mean $\lambda$.
\end{theorem}
\pause
\begin{align}
\pr(N_n=0) = \pr(t_{cov} \leq r) &\rightarrow e^{-e^{-x}} \nonumber \\
\pr(t_{cov} - n \log n\leq nx) &\rightarrow e^{-e^{-x}} \nonumber
\end{align}
\end{frame}

\begin{frame}{Birthday Problem}
What is the probability all birthdays are covered in a town of 1825?
How about a town of 2190?
\pause
\begin{align}
\pr(t_{cov} \leq 1825) &= \pr \left(\frac{t_{cov} - 2153}{365} \leq
\frac{328}{365}\right) \nonumber \\
&\approx e^{-e^{-0.89863}} \approx 0.08576. \nonumber
\only<3->{
\\\pr(t_{cov} \leq 2190) &= \pr \left(\frac{t_{cov} - 2153}{365} \leq
\frac{37}{365}\right) \nonumber \\
&\approx e^{-e^{-0.10137}} \approx 0.40511 \nonumber
}
\end{align}

@@

\begin{frame}{Birthday Problem}
\begin{tikzpicture}[scale=.37]
\node (graph) at (0,0) {\includegraphics[scale=.1, height=7cm]{birthdays}};
\node (title) at (1, 10) {Cover Times of the Birthday Problem (1000 trials)};
\node[rotate=90] (ylabel) at (-13,2) {$\pr(t_{cov} \leq r)$};
\node (xlabel) at (0,-9) {$r$};
\end{tikzpicture}
\end{frame}

\begin{frame}{Thank you!}
\begin{thebibliography}{9}

\bibitem{AF14}
	Aldous, D. and J. Fill.
	2014.
	\textit{Reversible Markov Chains and Random Walks on Graphs}.

\bibitem{BH94}
	Blom, G., L. Holst and D. Sandell.
	\textit{Problems and Snapshots from the World of Probability},
	Springer Science \& Business Media, 1994.

\bibitem{DS84}
	Doyle, P. G. and J. L. Snell.
	1984.
	\textit{Random Walks and Electric Networks},
	Mathematical Association of America.

\bibitem{LP08}
	Levin, D. A. and Y. Peres.
	2017.
	\textit{Markov Chains and Mixing Times},
	Vol. 107,
	American Mathematical Society.

\end{thebibliography}
\end{frame}
